---
title: "Untitled"
author: "v m kishore"
date: "03 Feb 2018"
output: html_document
---
#Reference 
#http://docs.h2o.ai/
#https://www.analyticsvidhya.com/blog/2016/05/h2o-data-table-build-models-large-data-sets/


```{r}
path <- "C:/Users/v m kishore/OneDrive/Data sets/AV/black friday/h20"
setwd(path)

#install and load the package
install.packages("data.table")
library(data.table)

#load data using fread
train <- fread("C:/Users/v m kishore/OneDrive/Data sets/AV/black friday/h20/train.csv", stringsAsFactors = T)
test <- fread("C:/Users/v m kishore/OneDrive/Data sets/AV/black friday/h20/test.csv", stringsAsFactors = T)


```


```{r}
#No. of rows and columns in Train
dim(train)
```

```{r}
#No. of rows and columns in Test
dim(test)
```


```{r}
str(train)
```

```{r}
#first prediction using mean
sub_mean <- data.frame(User_ID = test$User_ID, Product_ID = test$Product_ID, Purchase = mean(train$Purchase))
write.csv(sub_mean, file = "first_sub.csv", row.names = F)
```


```{r}
summary (train)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


```{r}
summary (test)
```

#combine data set
```{r}
test[,Purchase := mean(train$Purchase)]
c <- list(train, test)
combin <- rbindlist(c)
```


#analyzing gender variable
```{r}
combin[,prop.table(table(Gender))]
```


```{r}
combin[,prop.table(table(Occupation))]
```

```{r}
combin[,prop.table(table(Product_Category_1))]
```

```{r}
combin[,prop.table(table(Marital_Status))]
```

```{r}
length(unique(combin$Product_ID))
```

```{r}
length(unique(combin$User_ID))
```

```{r}
colSums(is.na(combin))
```


```{r}
library(ggplot2)
ggplot(combin, aes(Age, fill = Gender)) + geom_bar()
```


```{r}
ggplot(combin, aes(Age, fill = City_Category)) + geom_bar()
```




```{r}
install.packages('ggmosaic')
library(ggmosaic)
```

```{r}
ggplot(data = combin) +geom_mosaic(aes(x = product(Age), fill=City_Category))
```


```{r}
ggplot(data = combin) +geom_mosaic(aes(x = product(Age), fill=Gender))
```



```{r}
ggplot(combin, aes(x=Purchase,fill=Gender)) +
  geom_histogram()
```



```{r}
ggplot(aes(x = Purchase), data = combin) + 
  geom_freqpoly(aes(color = Gender), binwidth=10) + 
  scale_x_continuous(limits = c(0, 2400), breaks = seq(0, 2400, 50)) + 
  xlab('Purchase') + 
  ylab('Percentage of users shopping')
```



#Data Manipulation using data.table
```{r}
combin[,Product_Category_2_NA := ifelse(sapply(combin$Product_Category_2, is.na) ==    TRUE,1,0)]
combin[,Product_Category_3_NA := ifelse(sapply(combin$Product_Category_3, is.na) ==  TRUE,1,0)]
```

#impute missing values
```{r}
combin[,Product_Category_2 := ifelse(is.na(Product_Category_2) == TRUE, "-999",  Product_Category_2)]
combin[,Product_Category_3 := ifelse(is.na(Product_Category_3) == TRUE, "-999",  Product_Category_3)]
```


#set column level
```{r}
levels(combin$Stay_In_Current_City_Years)[levels(combin$Stay_In_Current_City_Years) ==  "4+"] <- "4"
```

#recoding age groups
```{r}
levels(combin$Age)[levels(combin$Age) == "0-17"] <- 0
levels(combin$Age)[levels(combin$Age) == "18-25"] <- 1
levels(combin$Age)[levels(combin$Age) == "26-35"] <- 2
levels(combin$Age)[levels(combin$Age) == "36-45"] <- 3
levels(combin$Age)[levels(combin$Age) == "46-50"] <- 4
levels(combin$Age)[levels(combin$Age) == "51-55"] <- 5
levels(combin$Age)[levels(combin$Age) == "55+"] <- 6
```

#convert age to numeric

```{r}
combin$Age <- as.numeric(combin$Age)
```

#convert Gender into numeric
```{r}
combin[, Gender := as.numeric(as.factor(Gender)) - 1]
```

#User Count
```{r}
combin[, User_Count := .N, by = User_ID]
```

#Product Count
```{r}
combin[, Product_Count := .N, by = Product_ID]
```

#Mean Purchase of Product
```{r}
combin[, Mean_Purchase_Product := mean(Purchase), by = Product_ID]
```

#Mean Purchase of User
```{r}
combin[, Mean_Purchase_User := mean(Purchase), by = User_ID]
```


```{r}
#one Hot Encoding
install.packages('dummies')
library(dummies)
combin <- dummy.data.frame(combin, names = c("City_Category"), sep = "_")
```


```{r}
#Divide into train and test
c.train <- combin[1:nrow(train),]
c.test <- combin[-(1:nrow(train)),]
```

```{r}
c.train <- c.train[c.train$Product_Category_1 <= 18,]
```

#install JDK
#http://www.oracle.com/technetwork/java/javase/downloads/jdk10-downloads-4416644.html
```{r}
install.packages("h2o")
library(h2o)
```


```{r}
localH2O <- h2o.init(nthreads = -1)
```

```{r}
h2o.init()
```

#data to h2o cluster
```{r}
train.h2o <- as.h2o(c.train)
test.h2o <- as.h2o(c.test)
```

#dependent variable (Purchase)
```{r}
y.dep <- 14
x.indep <- c(3:13,15:20)
```

#Random Fores
```{r}
system.time(rforest.model <- h2o.randomForest(y=y.dep, x=x.indep, training_frame = train.h2o, ntrees = 1000, mtries = 3, max_depth = 4, seed = 1122))
```


```{r}
h2o.performance(rforest.model)
```


```{r}
h2o.varimp(rforest.model)
```

```{r}
h2o.shutdown()
```

