---
title: "Hypothesis-Test"
author: "V M KISHORE"
date: "2/21/2017"
output:
  pdf_document: default
  html_document: default
---

Referance 
https://www.r-bloggers.com/two-sample-students-t-test-1/

F test to compare two variances
```{r}
a = c(175, 168, 168, 190, 156, 181, 182, 175, 174, 179)
b = c(185, 169, 173, 173, 188, 186, 175, 174, 179, 180)

Ftest<-var.test(a,b)
Ftest
```




#Difference between means Hypothesis test usnig t distribution(studient distribution)
```{r}
t.test(a,b, var.equal=TRUE, paired=FALSE)
```




#Chi-Square Test(Relation between tow categorical variables)
  -Chi-Square test is a statistical method to determine if two categorical
  -variables have a significant correlation between them
  
References
http://stattrek.com/chi-square-test/goodness-of-fit.aspx?Tutorial=AP
http://www.statisticshowto.com/what-is-a-chi-square-statistic/
http://www.r-tutor.com/elementary-statistics/goodness-fit/chi-squared-test-independence


Load the library.
```{r}
library("MASS")

```


Create a data frame from the main data set.
```{r}
car.data <- data.frame(Cars93$AirBags, Cars93$Type)

```


Create a table with the needed variables.
```{r}
car.data = table(Cars93$AirBags, Cars93$Type)
print(car.data)

```

```{r}
print(chisq.test(car.data))

```





*Analysis of covariance (categorical variable VS continues.Analysis of Covariance also called as ANCOVA.)
Model with interaction between categorical variable and predictor variable
Get the dataset.
```{r}
input <- mtcars

```


Create the regression model.
```{r}
result <- aov(mpg~hp*am,data = input)
print(summary(result))
```



Model without interaction between categorical variable
Get the dataset.
```{r}
input <- mtcars

```


Create the regression model.
```{r}
result <- aov(mpg~hp+am,data = input)
print(summary(result))
```



Comparing Two Models
Get the dataset.
```{r}
input <- mtcars

```



Create the regression models.
```{r}
result1 <- aov(mpg~hp*am,data = input)
result2 <- aov(mpg~hp+am,data = input)

```


Compare the two models.
```{r}
print(anova(result1,result2))
```

*ANOVA test knowing P value
```{r}
pf(21.5,3,791,lower.tail=FALSE)
```



#Paul inferance test(Inferance for small proportions)
#source("http://bit.ly/dasi_inference")
```{r}
paul=factor(c(rep("yes",8),rep("no",0)),level=c("yes","no"))
install.packages('inference')
library(inference)
inference(paul,est="proportion",type='ht',method="simulation",success="yes",null=.5,alternative="greater")

```

